---
title: "Explora_Data_Survey"
author: "Mauricio Mardones I."
date: "9/23/2021"
output: pdf_document
toc: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r lib, warning=F, message=F, error=F}
library(GGally)
library(tidyverse, quietly = TRUE)
library(patchwork)
library(marmap)
library(ggplot2)
library(raster)
library(dbplyr)
library(knitr)
library(data.table)
```


```{r data2}
# Genero una carpeta en donde alojar figuras
dir.Fig        <-"Figuras/"
fig            <-c("pdf","bmp")
```


# Base de de los survey

Base de datos provista en WG-EMM-2021/05 Rev. . El codigo es un provisto por Tracey Dorna
### Set up document, files and packages

Packages used: 
 
    Matt Dowle and Arun Srinivasan (2020). data.table: Extension of `data.frame`. R package version 1.13.4.
    https://CRAN.R-project.org/package=data.table  
   
    Hadley Wickham (2016). ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York.  
   
    Yihui Xie (2020). knitr: A General-Purpose Package for Dynamic Report Generation in R. R package version 1.30.   

   

  

### Read in data 

````{r}
mdat <- read.csv("ASAM_metadata_2021_v2_tidy_0.csv")
```

### Tidy and inspect file structure

  1. Checking structure (most data is currently character type)
  2. Renaming columns to be R friendly
  3. Inspect individual column contents for unusual formats or notes prior to type conversion
  

### Names of Data:

Set names to be more R Friendly 

```{r data inspection, include=TRUE}
#names(mdat)
# print table of metadata column names and R coding names
knitr::kable(as.data.frame(cbind(`Metadata Table Names`=names(mdat), 
                    `R Names`=c("Year_yyyy",
                      "Month_MON",
                      "Vessel",
                      "Contributor",
                      "Subarea",
                      "Survey_name",
                      "Density_gm2",
                      "CV_of_density_Perc",
                      "CV_method",
                      "Survey_area_km2",
                      "Echosounder",
                      "Freq_for_biomass_est_kHz",
                      "Frequencies_avail",
                      "TS_Id_Method",
                      "dB_diff_window",
                      "TS_model",
                      "Depth_range_integrated_m",
                      "Time_sampled",
                      "Stratum_name",
                      "Survey_design_description", 
                      "Reference", 
                      "Note", 
                      "empty1", 
                      "sourceexl",
                      "Net", 
                      "Towdesign", 
                      "ASAM_NOTES"))))

#rename columns to be code friendly
setnames(mdat, names(mdat), c("Year_yyyy",
                              "Month_MON",
                              "Vessel",
                              "Contributor",
                              "Subarea",
                              "Survey_name",
                              "Density_gm2",
                              "CV_of_density_Perc",
                              "CV_method",
                              "Survey_area_km2",
                              "Echosounder",
                              "Freq_for_biomass_est_kHz",
                              "Frequencies_avail",
                              "TS_Id_Method",
                              "dB_diff_window",
                              "TS_model",
                              "Depth_range_integrated_m",
                              "Time_sampled",
                              "Stratum_name",
                              "Survey_design_description", 
                              "Reference", 
                              "Note", 
                              "empty1", 
                              "sourceexl",
                              "Net", 
                              "Towdesign", 
                              "ASAM_NOTES"))

```

### Years & Months available:

```{r yrs, include=TRUE}
sort(unique(mdat$Year_yyyy))

unique(mdat$Month_MON) 

```

### Vessels and contributors

```{r VnCs, include=TRUE}
unique(mdat$Vessel)
unique(mdat$Contributor)   
```

### Subarea codes
 
Area "48" is the ccamlr 2000 survey
```{r subareas, include=TRUE}                                
unique(mdat$Subarea) # Area "48" is the ccamlr 2000 survey

```

### CV method


```{r cvmeths, include=TRUE}
unique(mdat$CV_method) 
```
  
There is an unusual entry of CV method:

  "Here, the CV were simply calculated as the S.E/Mean x 100% for each stratum or entire survey area"
  
A note has been added to the "Note" column indicating that "CV calculated as the S.E/Mean x 100% for each stratum or entire survey area"  


 **Data where CV calculated as the S.E/Mean x 100% for each stratum or entire survey area**
 
```{r cv 95 CIs, include=TRUE}
# INSPECT and add to ASAM_NOTES column
knitr::kable(mdat[mdat$CV_method=="Here, the CV were simply calculated as the S.E/Mean x 100% for each stratum or entire survey area", c(1:8, 10)])
```

### Echosounder used:

```{r echos, include=TRUE}
unique(mdat$Echosounder)

```

### Frequency used for biomass estimation, TS method and TS model

```{r Freq TSs, include=TRUE}
unique(mdat$Freq_for_biomass_est_kHz)
unique(mdat$TS_Id_Method)
unique(mdat$TS_model)

```

### Depth range integrated

```{r dint, include=TRUE}
unique(mdat$Depth_range_integrated_m)

```

### Time of samples

```{r times, include=TRUE}
unique(mdat$Time_sampled) # ? should we consolidate "daylight only" to be "day"

```

### Stratum names and codes

To ease coding new strata codes were set up for Area 48.1 data under col 'strata'.  

Strata that were assigned a code based on the area they matched best.  

Surveys which overlapped multiple areas were assigned combined codes.

```{r strata_filter, echo=TRUE}
unique(mdat$Stratum_name)
```

### Survey design

```{r survey desc, include=TRUE}

unique(mdat$Survey_design_description)

```

### Notes/Comments

```{r note, include=TRUE}
unique(mdat$Note)
unique(mdat$empty1)
unique(mdat$ASAM_NOTES)

#mdat[mdat$ASAM_NOTES=="", mdat$ASAM_NOTES=="NA"] # getting rid of empty ""

```

### Survey details

Tow design. 
Nets.
Source exl. 

```{r details, include=TRUE}

unique(mdat$Towdesign)

unique(mdat$Net)

unique(mdat$sourceexl)
```


### Remove Duplicate and Bad data

As analysis requires Density CV and Area for weighted density calculations:
 
  1.	Remove rows which do not have complete records for ‘Density’ and/or ‘CV’ 
  2.	Remove anything with a comment in the ‘ASAM_NOTES’ because this was either
    a.	the same AMLR data but run with the Greene algorithm so DUPLICATED 
    b.	incomplete/the area wasn’t covered properly so difficult to weight appropriately

**Removed data is being saved in a table named remdat**

```{r remove dup}

# Retain a data.table with data removed
#remdat <- mdat[is.na(mdat$ASAM_NOTES)]

# Remove duplicated or other data with comments in ASAM_NOTES
#mdat <- mdat[is.na(mdat$ASAM_NOTES)]

```
 
### Density 

Miramos los datos de densidades


```{r inspect dens, include=TRUE}
unique(mdat$Density_gm2)
```


```{r hdba, warning=F, include=T, message=F, echo=FALSE,fig.height=7,fig.width=9,fig.align="center",fig.path=dir.Fig,dev=fig, fig.cap="Histo Dens by Area"}
mdat$Density_gm2<- as.double(mdat$Density_gm2)
h <- ggplot(mdat, aes(x=as.double(Density_gm2)))+
  geom_histogram(binwidth=5, show.legend = FALSE, fill=2)+
  facet_wrap(~Subarea, ncol =3)+
  theme_bw()
h
```

A plot with density in grs.


```{r densidad by, warning=F, include=T, message=F, echo=FALSE,fig.height=6,fig.width=8,fig.align="center",fig.path=dir.Fig,dev=fig, fig.cap="Densidad by subArea"}
sc <- ggplot(mdat, aes(x=Year_yyyy, y=as.double(mdat$Density_gm2))) + 
        geom_point() +
        geom_smooth(method = "loess", colour='red', fill='red', alpha=.3)+
        scale_x_continuous(breaks = seq(from = 190, to = 2019, by = 4))+
        theme_bw()+
        theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=9))+
        theme(axis.text.x = element_text(angle = 90, hjust = 2))+
        facet_wrap(~Subarea, ncol =3)+
        ggtitle('Density Krill Survey (gr.)')+
        ylim(0,400)+
        xlim(1995,2020)
sc

```
 
#### CV

Some CV values are 95% CI range rather than actual CV.

Remove from analysis data set (store in remdat).

```{r inspect CV, include=TRUE}

unique(mdat$CV_of_density_Perc)

mdat$CV_of_density_Perc <- as.double(mdat$CV_of_density_Perc)
hist(mdat$CV_of_density_Perc, breaks = 100)

# 
# knitr::kable(mdat[CV_of_density_Perc=="95%CI 0-6.42"|CV_of_density_Perc=="95%CI 0.21-11.29", 1:10])
# 
# # The above do not have CV values we can work with so removing them from analysis
# remdat <- rbind(remdat,  mdat[CV_of_density_Perc %in% c("95%CI 0-6.42", "95%CI 0.21-11.29")])
# mdat <- mdat[!CV_of_density_Perc %in% c("95%CI 0-6.42", "95%CI 0.21-11.29")]

```
