---
title: "Data estimates biomass AMLR data by "
subtitle: "Complemetnary analysis to incorporate in Krill Stock Assessment Model 48.1 SubArea"
author: "Mardones, M; Watters, G.; Cárdenas, C."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: biomass.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup1}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      dpi = 300)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```



```{r lib, warning=F, message=F, error=F}
library(GGally)
library(tidyverse, quietly = TRUE)
library(patchwork)
library(marmap)
library(ggplot2)
library(raster)
library(dbplyr)
library(knitr)
library(data.table)
library(ggthemes)
library(egg)
library(ggrepel)
```


```{r data2}
# Genero una carpeta en donde alojar figuras
dir.Fig        <-"Figuras/"
fig            <-c("pdf","bmp")
```


# BACKGROUND

The following document intends to carry out a complementary
methodological analysis to explore with the
population dynamics of krill (*Euphausia superba*), in this case, with a index of biommass based in AMLR Program Survey with both standarized and calibrated data.


# OBJECTIVE

Once the correlation and effects on the population and/or fishing
indicators on krill have been verified, this analysis aims to have a
time series of the environmental variable to incorporate into the stock assessment process. Similar work in @Wang2021 but with a longest fishery history and adding environmental and predator data.


# METODOLOGY

### Read data and code provided by  WG-EMM-2021/05 Rev.
Different packages was used  in this report like @Yuhui2023, @Ludecke2021, @Report2023. 

```{r warning=FALSE}
# Define preferred figure format
My_theme <- theme(axis.text.y = element_blank(),
                  axis.ticks.y = element_blank(),
                  axis.ticks.x=element_blank(),
                  panel.background = element_blank(),
                  panel.border = element_rect(fill = NA, 
                                              size =  1),
                  strip.background = element_rect(fill = "white", 
                                                  color = "white", 
                                                  size = 1),
                  text = element_text(size = 14),
                  panel.grid.major = element_line(colour = "white", 
                                                  size = 0.1),
                  panel.grid.minor = element_line(colour = "white", 
                                                  size = 0.1))
```


```{r}
mdat <- read.csv("ASAM_metadata_2021_v2_tidy_0.csv")
names(mdat)
```

### Tidy and inspect file structure

  1. Checking structure (most data is currently character type)
  2. Renaming columns to be R friendly
  3. Inspect individual column contents for unusual formats or notes prior to type conversion
  

### Names of Data

Set names to be more R Friendly 

```{r data inspection, include=TRUE, echo=FALSE}
#names(mdat)
# print table of metadata column names and R coding names
knitr::kable(as.data.frame(cbind(`Metadata Table Names`=names(mdat), 
                    `R Names`=c("Year_yyyy",
                      "Month_MON",
                      "Vessel",
                      "Contributor",
                      "Subarea",
                      "Survey_name",
                      "Density_gm2",
                      "CV_of_density_Perc",
                      "CV_method",
                      "Survey_area_km2",
                      "Echosounder",
                      "Freq_for_biomass_est_kHz",
                      "Frequencies_avail",
                      "TS_Id_Method",
                      "dB_diff_window",
                      "TS_model",
                      "Depth_range_integrated_m",
                      "Time_sampled",
                      "Stratum_name",
                      "Survey_design_description", 
                      "Reference", 
                      "Note", 
                      "empty1", 
                      "sourceexl",
                      "Net", 
                      "Towdesign", 
                      "ASAM_NOTES"))))

#rename columns to be code friendly
setnames(mdat, names(mdat), c("Year_yyyy",
                              "Month_MON",
                              "Vessel",
                              "Contributor",
                              "Subarea",
                              "Survey_name",
                              "Density_gm2",
                              "CV_of_density_Perc",
                              "CV_method",
                              "Survey_area_km2",
                              "Echosounder",
                              "Freq_for_biomass_est_kHz",
                              "Frequencies_avail",
                              "TS_Id_Method",
                              "dB_diff_window",
                              "TS_model",
                              "Depth_range_integrated_m",
                              "Time_sampled",
                              "Stratum_name",
                              "Survey_design_description", 
                              "Reference", 
                              "Note", 
                              "empty1", 
                              "sourceexl",
                              "Net", 
                              "Towdesign", 
                              "ASAM_NOTES"))

```

### Years & Months available:

```{r yrs, include=TRUE}
sort(unique(mdat$Year_yyyy))

unique(mdat$Month_MON) 

```

### Vessels and contributors

```{r VnCs, include=TRUE}
unique(mdat$Vessel)
unique(mdat$Contributor)   
```

### Subarea codes
 
Area "48" is the ccamlr 2000 survey and strata
```{r subareas, include=TRUE}                                
unique(mdat$Subarea) # Area "48" is the ccamlr 2000 survey
unique(mdat$Stratum_name)
```

### CV method


```{r cvmeths, include=TRUE}
unique(mdat$CV_method) 
```
  
There is an unusual entry of CV method:

  "Here, the CV were simply calculated as the S.E/Mean x 100% for each stratum or entire survey area"
  
A note has been added to the "Note" column indicating that "CV calculated as the S.E/Mean x 100% for each stratum or entire survey area"  


 **Data where CV calculated as the S.E/Mean x 100% for each stratum or entire survey area**
 
```{r cv 95 CIs, include=TRUE}
# INSPECT and add to ASAM_NOTES column
knitr::kable(mdat[mdat$CV_method=="Here, the CV were simply calculated as the S.E/Mean x 100% for each stratum or entire survey area", c(1:8, 10)])
```

### Echosounder used:

```{r echos, include=TRUE}
unique(mdat$Echosounder)

```

### Frequency used for biomass estimation, TS method and TS model

```{r Freq TSs, include=TRUE}
unique(mdat$Freq_for_biomass_est_kHz)
unique(mdat$TS_Id_Method)
unique(mdat$TS_model)

```

### Depth range integrated

```{r dint, include=TRUE}
unique(mdat$Depth_range_integrated_m)

```

### Time of samples

```{r times, include=TRUE}
unique(mdat$Time_sampled) # ? should we consolidate "daylight only" to be "day"

```

### Stratum names and codes

To ease coding new strata codes were set up for Area 48.1 data under col 'strata'.  

Strata that were assigned a code based on the area they matched best.  

Surveys which overlapped multiple areas were assigned combined codes.

```{r echo=TRUE}
unique(mdat$Stratum_name)
```

### Survey design

```{r survey desc, include=TRUE}

unique(mdat$Survey_design_description)

```

### Notes/Comments

```{r note, include=TRUE}
unique(mdat$Note)
unique(mdat$empty1)
unique(mdat$ASAM_NOTES)

#mdat[mdat$ASAM_NOTES=="", mdat$ASAM_NOTES=="NA"] # getting rid of empty ""

```

### Survey details

Tow design. 
Nets.
Source exl. 

```{r details, include=TRUE}

unique(mdat$Towdesign)

unique(mdat$Net)

unique(mdat$sourceexl)
```


### Remove Duplicate and Bad data

As analysis requires Density CV and Area for weighted density calculations:
 
  1.	Remove rows which do not have complete records for ‘Density’ and/or ‘CV’ 
  2.	Remove anything with a comment in the ‘ASAM_NOTES’ because this was either
    a.	the same AMLR data but run with the Greene algorithm so DUPLICATED 
    b.	incomplete/the area wasn’t covered properly so difficult to weight appropriately

**Removed data is being saved in a table named remdat**

```{r remove dup}

# Retain a data.table with data removed
#remdat <- mdat[is.na(mdat$ASAM_NOTES)]

# Remove duplicated or other data with comments in ASAM_NOTES
#mdat <- mdat[is.na(mdat$ASAM_NOTES)]

```
 
### Density 

Take look densities

```{r inspect dens, include=TRUE}
unique(mdat$Density_gm2)
```


```{r hdba}
mdat$Density_gm2<- as.double(mdat$Density_gm2)
h <- ggplot(mdat, aes(x=as.double(Density_gm2)))+
  geom_histogram(binwidth=5, show.legend = FALSE, fill=2)+
  facet_wrap(~Stratum_name, ncol =6)+
  theme_bw()
h
```

A plot with density in grs. by stratum


```{r densidad byor, warning=FALSE}
sc <- ggplot(mdat, aes(x=Year_yyyy, y=as.double(mdat$Density_gm2))) + 
        geom_point() +
        geom_smooth(method = "loess", colour='red', fill='red', alpha=.3)+
        scale_x_continuous(breaks = seq(from = 190, to = 2019, by = 4))+
        theme_bw()+
        theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=9))+
        theme(axis.text.x = element_text(angle = 90, hjust = 2))+
        facet_wrap(~Stratum_name, ncol =6)+
        ggtitle('Density Krill Survey (gr.)')+
        ylim(0,400)+
        xlim(1995,2020)
sc

```
 
### CV

Some CV values are 95% CI range rather than actual CV.

Remove from analysis data set (store in remdat).

```{r inspect CV, include=TRUE}

unique(mdat$CV_of_density_Perc)

mdat$CV_of_density_Perc <- as.double(mdat$CV_of_density_Perc)
hist(mdat$CV_of_density_Perc, breaks = 100)

# 
# knitr::kable(mdat[CV_of_density_Perc=="95%CI 0-6.42"|CV_of_density_Perc=="95%CI 0.21-11.29", 1:10])
# 
# # The above do not have CV values we can work with so removing them from analysis
# remdat <- rbind(remdat,  mdat[CV_of_density_Perc %in% c("95%CI 0-6.42", "95%CI 0.21-11.29")])
# mdat <- mdat[!CV_of_density_Perc %in% c("95%CI 0-6.42", "95%CI 0.21-11.29")]

```
### Stratum names and codes

To ease coding new strata codes were set up for Area 48.1 data under col 'strata'.  

Strata that were assigned a code based on the area they matched best.  

Surveys which overlapped multiple areas were assigned combined codes.

```{r echo=TRUE}
unique(mdat$Stratum_name)
```


```{r echo=TRUE}
mdat$strata[mdat$Stratum_name%in%c("Elephant Island", "Elephant Island (EL)")] <- "E"    
mdat$strata[mdat$Stratum_name%in%c("Joinville")] <- "J"    
mdat$strata[mdat$Stratum_name%in%c("South", "Bransfield",  "Bransfield Strait (BS)")] <- "S"    
mdat$strata[mdat$Stratum_name%in%c("West", "South Shetland Islands North", "South Shetland Island (SSI)")] <- "W"   
mdat$strata[mdat$Stratum_name%in%c("ESS", "entire survey area")] <- "WESJ"   
mdat$strata[mdat$Stratum_name%in%c("AP")] <- "WE"   
mdat$strata[mdat$Stratum_name%in%c("South Shetland Islands")] <- "WS"   
mdat$strata[mdat$Stratum_name%in%c("SSI")] <- "With_AP_is_WESJ"   
table(mdat$strata)
# test NA Data
colSums(is.na(mdat))

# drop NA in Strata
mdat1 <- mdat %>% 
  drop_na()
```

```{r densidad by strata, warning=F}
destra <- ggplot(mdat1, 
                 aes(x=Year_yyyy, 
                     y=as.double(mdat1$Density_gm2),
                     color=TS_Id_Method)) +
  geom_point() +
  geom_smooth(method = "loess", colour='grey', fill='grey', alpha=.3)+
  scale_x_continuous(breaks = seq(from = 1990, to = 2020, by = 2))+
  scale_color_viridis_d(option = "E")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=9),
        panel.grid.major = element_line(colour = "white", size = 0.1),
        panel.grid.minor = element_line(colour = "white", size = 0.1))+
  facet_wrap(~strata, ncol =3)+
  ggtitle('Density Krill Survey (gr.)')+
  ylim(0,200)+
  labs(y="Density (gr.)",
       x= "")
destra
```
```{r densid, warning=F}
destra <- ggplot(mdat1, 
                 aes(x=Year_yyyy, 
                     y=as.double(mdat1$Density_gm2),
                     color=Depth_range_integrated_m)) +
  geom_point() +
  geom_smooth(method = "loess", 
              colour='grey', 
              fill='grey', alpha=.3)+
  scale_x_continuous(breaks = seq(from = 1990, to = 2020, by = 2))+
  scale_color_viridis_d(option = "D",
                        name="Deep Range")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=9),
        panel.grid.major = element_line(colour = "white", size = 0.1),
        panel.grid.minor = element_line(colour = "white", size = 0.1))+
  facet_wrap(~strata, ncol =3)+
  ggtitle('Density Krill Survey (gr.)')+
  ylim(0,200)+
  labs(y="Density (gr.)",
       x= "")
destra
```
Some statistics descriptive

```{r}
library(pastecs)
descrs <- stat.desc(mdat[,c(7,8, 12)])
kable(descrs,
      format="html",
      digits = 3)

skimr::skim(mdat)
```
In this part we star to work with procesed data from WG-EEM-21

# Biomass Depured Data
Load Biomass depured data


```{r}
bioma <- read_csv("krill_biomass_acoustic_survey.csv")
View(bioma)

```
Some statistics to biomass calculation


```{r}
descrsbio <- stat.desc(bioma[,c(-1)])
kable(descrsbio,
      format="html",
      digits = 3)

skimr::skim(bioma)

```


Plotr biomass calculation by Strata
```{r}
strbio <-ggplot(bioma%>% 
                 filter(Strata != "WestElephant(AP)"), 
                aes(Year,
                    stratabiomass,
               fill=Strata))+ 
  geom_col()+
  facet_wrap(.~Strata,
             ncol=1)+
  scale_fill_viridis_d(option="D")+
  scale_x_continuous(breaks = seq(from = 1990, to = 2020, by = 1))+
  scale_y_continuous(labels = scales::number_format()) +
  theme_few() +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=9),
        panel.grid.major = element_line(colour = "white", size = 0.1),
        panel.grid.minor = element_line(colour = "white", size = 0.1),
        legend.position = "none")+
  labs(y="Biomass (t)")
strbio
```

Another important parameters is el CV. 


```{r warning=FALSE}
cvbio <-ggplot(bioma %>% 
                 filter(Strata != "WestElephant(AP)"),
               aes(Year, cvbiomass,
               color=Strata))+ 
  geom_point()+
  facet_wrap(.~Strata,
             ncol=1)+
  scale_color_viridis_d(option="D")+
  scale_x_continuous(breaks = seq(from = 1990, 
                                  to = 2020, 
                                  by = 1))+
  geom_text_repel(aes(label = round(cvbiomass,1)),
                  size = 1.5,
                  color="black")+
  theme_few() +
  theme(axis.text.x  = element_text(angle=90, 
                                    vjust=0.5, 
                                    size=9),
        panel.grid.major = element_line(colour = "white", 
                                        size = 0.1),
        panel.grid.minor = element_line(colour = "white", 
                                        size = 0.1),
        legend.position = "none")+
  labs(y="CV (%) Survey")+
  ylim(0,100)
cvbio

```

JOin maps

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=4}
ggarrange(strbio, cvbio, ncol = 2)
```

Sum of biomass considering strata and year.

```{r}
biomasto <- bioma %>% 
  group_by(Strata, Year) %>% 
  dplyr::summarise(biot=sum(stratabiomass),
                   cvto=mean(cvbiomass))
```
write a `csv` output with biomass and CV

```{r eval=FALSE}
write_csv(biomasto, "biomassurvey.csv")
```
 
 last plot with total

```{r  out.width = '60%'}
# plot
biot <- ggplot(biomasto %>% 
                 group_by(Year) %>% 
                 summarise(biototal =sum(biot)),
               aes(Year, biototal))+
  geom_point(shape=21,
             size=3,)+
  geom_text_repel(aes(label = biototal),
                  size = 3,
                  color="black")+
  geom_smooth(method="lm",
              color="red")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=9),
        panel.grid.major = element_line(colour = "white", size = 0.1),
        panel.grid.minor = element_line(colour = "white", size = 0.1))+
  scale_x_continuous(breaks = seq(from = 1990, to = 2020, by = 1))+
  labs(y="Biomass Survey (t)",
       x="")
biot
```


# NEXT STEP

- Biomass vector to add in stock assessment model as index relative about biomass with Catchability. Catchability patterns indicate that the catchability coefficient has been used in two main lines: (a) increased efficiency of fishing effort and (b) its relation to population fishery processes for assessment and management purposes. It involves various aspects of the fishery, such as individual and population biology, characteristics of the fishing gear, amount of fishing, fishing strategies, and environmental fluctuation, among others-

- Another component relevant is CV, because this uncertainty must  be add in stock assesment.

# REFERENCES

